Extended Data

### 2022-03-08 doc/manuscript/Dow_manuscript_Nature.Rmd
Update Dow_manuscript_Nature.Rmd
@teixeirak I've mad some slight changes to your new intro. Feel free to revert as you like

In temperate deciduous forests, spring warming generally lengthens the period over which trees have photosynthetically active leaves[@jeong_phenology_2011;@ibanez_forecasting_2010;@menzel_european_2006;@menzel_growing_1999;@lu_no_2022] and that over which the ecosystem is a net CO~2~ sink[@keenan_net_2014;@lu_no_2022].
Based on these observations, current terrestrial ecosystem models effectively assume that longer growing seasons lead to sustained increases in annual net CO~2~ uptake (i.e., net ecosystem exchange, _NEE_)[@buermann_widespread_2018;@zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012].
However, model representation of C allocation to stem growth -- or woody aboveground net primary productivity ($ANPP_{woody}$) on the ecosystem level -- and its seasonality is simplistic relative to *known decoupling of photosynthate production and stem growth [@delpierre_temperate_2016;@delpierre_wood_2016]*. <!-- go back and check that these references are appropriate for this statement-->
Given that the C allocated to woody growth has the longest residence time in the ecosystem[@refs], understanding whether warming spring temperatures will lead to sustained increases in forest C sequestration requires understanding the sensitivity of $ANPP_{woody}$ to spring temperatures.

We have little evidence as to how stem growth and $ANPP_{woody}$ respond to warmer spring temperatures in deciduous forests[@savage_coordination_2021;@delpierre_temperate_2016]. 
Most studies on tree growth responses to warmer spring temperatures have focused on boreal or temperate conifers, which tend to respond to warmer spring temperatures with an earlier start to growth[@huang_photoperiod_2020;@delpierre_chilling_2019;@rossi_critical_2008] and increased annual growth in mesic climates[@gao_earlier_2022;@anderson-teixeira_joint_2021; @more_conifer_refs]. 
Much less is known about deciduous tree responses[@savage_coordination_2021;@delpierre_temperate_2016].
Close coordination of budburst and initiation of xylogenesis[@delpierre_temperate_2016] suggests that warm springs should shift the onset of growth earlier alongside observed advances in leaf phenology[@jeong_phenology_2011;@keenan_net_2014;@buermann_widespread_2018;@finzi_carbon_2020;@stuble_plant_2021;@delgado_differences_2020].
However, earlier growth would not necessarily translate to earlier, faster, or more stem growth; rather, cell enlargement is dependent upon environmental conditions on hourly to daily time scales[@zweifel_why_2021], and annual growth is more closely linked to conditions within the growing season than to growing season length[@etzold_number_2022]. 
Furthermore, *growth of broadleaf deciduous trees may be sink-saturated*, such that longer growing seasons with more C fixation do not necessarily augment growth[@refs].
Tree-ring records, which can be used to examine relationships of annual growth to temperature but not to understand growth seasonality, reveal that growth of temperate deciduous trees tends to be most sensitive to temperature or potential evapotranspiration between late spring and early summer[@dorangeville_drought_2018;@helcoski_growing_2019], with some evidence that warmer springs may have a modest, but usually insignificant, positive effect on growth[@helcoski_growing_2019;@gao_earlier_2022].
These observations do not necessarily align with the finding that warming spring temperatures increase annual forest CO~2~ uptake in temperate deciduous forests[@keenan_net_2014;@lu_no_2022].
Characterizing responses of stem growth to warming spring temperatures is critical to bridging this conceptual disconnect and understanding how forest biomass growth is likely to change as the climate warms[@refs].

>>>>

In temperate deciduous forests, spring warming generally lengthens the period over which trees have photosynthetically active leaves[@jeong_phenology_2011;@ibanez_forecasting_2010;@menzel_european_2006;@menzel_growing_1999;@lu_no_2022] and that over which the ecosystem is a net CO~2~ sink[@keenan_net_2014;@lu_no_2022].
Based on these observations, current terrestrial ecosystem models effectively assume that longer growing seasons lead to sustained increases in annual net CO~2~ uptake (i.e., net ecosystem exchange, _NEE_)[@buermann_widespread_2018;@zani_increased_2020;@keenan_net_2014;@ahlstrom_robustness_2012].
#However, this assumption is overly simplistic given the known decoupling of photosynthate production and stem growth[@delpierre_temperate_2016;@delpierre_wood_2016], suggesting that model representaiton of C allocation to stem growth -- or woody aboveground net primary productivity ($ANPP_{woody}$) on the ecosystem level -- needs to be reworked.# 
However, model representation of C allocation to stem growth -- or woody aboveground net primary productivity ($ANPP_{woody}$) on the ecosystem level -- and its seasonality is simplistic relative to *known decoupling of photosynthate production and stem growth [@delpierre_temperate_2016;@delpierre_wood_2016]*. <!-- go back and check that these references are appropriate for this statement-->
Given that the C allocated to woody growth has the longest residence time in the ecosystem[@refs], understanding whether warming spring temperatures will lead to sustained increases in forest C sequestration requires understanding the sensitivity of $ANPP_{woody}$ to spring temperatures.

We have little evidence as to how stem growth and $ANPP_{woody}$ respond to warmer spring temperatures in deciduous forests[@savage_coordination_2021;@delpierre_temperate_2016]. 
Most studies on tree growth responses to warmer spring temperatures have focused on boreal or temperate conifers, which tend to respond to warmer spring temperatures with an earlier start to growth[@huang_photoperiod_2020;@delpierre_chilling_2019;@rossi_critical_2008] and increased annual growth in mesic climates[@gao_earlier_2022;@anderson-teixeira_joint_2021; @more_conifer_refs]. 
Close coordination of budburst and initiation of xylogenesis[@delpierre_temperate_2016] suggests that warm springs should shift the onset of growth earlier in deciduous trees alongside observed advances in leaf phenology[@jeong_phenology_2011;@keenan_net_2014;@buermann_widespread_2018;@finzi_carbon_2020;@stuble_plant_2021;@delgado_differences_2020].
However, an earlier start to growth would not necessarily translate to greater or faster stem growth; instead, growth is dependent upon environmental conditions on hourly to daily time scales[@zweifel_why_2021] within the growing season rather than to growing season length[@etzold_number_2022]. 
Furthermore, *growth of broadleaf deciduous trees may be sink-saturated*, resulting in the lack of more C fixation to stem growth in extended growing seasons[@refs].
Tree-ring records, which can be used to examine relationships of annual growth to temperature but not to understand growth seasonality, reveal that growth of temperate deciduous trees tends to be most sensitive to temperature or potential evapotranspiration between late spring and early summer[@dorangeville_drought_2018;@helcoski_growing_2019], with some evidence that warmer springs may have a modest, but usually insignificant, positive effect on growth[@helcoski_growing_2019;@gao_earlier_2022].
These observations do not necessarily align with the finding that warming spring temperatures increase annual forest CO~2~ uptake in temperate deciduous forests[@keenan_net_2014;@lu_no_2022].
Characterizing responses of stem growth to warming spring temperatures is critical to bridging this conceptual disconnect and understanding how forest biomass growth is likely to change as the climate warms[@refs].
